---
title: "Comprehensive Output Analysis Tools"
output: 
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: header.html
editor_options: 
  markdown: 
    wrap: 72
---

This vignette introduces how to use `plot_mse_output()` function to 
summarize and visualize the results of management strategy evaluation.
`plot_mse_output()` function is capable of summarizing management
strategy evaluation results from one realization and from n
realizations, when using parallel computing (see Example 5 for more
details)). The results can be presented in line plot, scatter plot, box
plot, and radar chart for users to better compare performance among
different management strategies. Please use ?plot_mse_output to see more
details. Below are one example about plotting results from one
realization and one example about plotting results from n realizations.

### 1. Generate operating model

```{r, eval=FALSE}
library(wham)
library(SPASAM.MSE)

main.dir = here::here() # path to save the results.

year_start  <- 1  # starting year in the burn-in period
year_end    <- 30  # end year in the burn-in period
MSE_years   <- 10     # number of years in the feedback loop
# Note: no need to include MSE_years in simulation-estimation 

info <- generate_basic_info(n_stocks = 2, 
                            n_regions = 2,
                            n_indices = 2,
                            n_fleets = 2,
                            n_seasons = 1,
                            base.years = year_start:year_end,
                            n_feedback_years = MSE_years,
                            life_history = "short",
                            n_ages = 10) 

basic_info = info$basic_info # collect basic information
catch_info = info$catch_info # collect fleet catch information
index_info = info$index_info # collect survey information
F_info = info$F # collect fishing information

# Create difference in NAA between 2 stocks
n_stocks   <- 2
n_regions  <- 2
n_ages     <- 10
sigma      <- "rec+1"
re_cor     <- "iid"
ini.opt    <- "equilibrium" # option   <- c("age-specific-fe", "equilibrium")

# Set sigma for NAA
NAA_sig <- 0.2
sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages"
sigma_vals[,,1] = 0.5

NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               recruit_model = 2,
               sigma_vals = sigma_vals)

# Prepare wham input
input <- prepare_wham_input(basic_info = basic_info, 
                            NAA_re = NAA_re,
                            catch_info = catch_info, 
                            index_info = index_info, 
                            F = F_info)

random = input$random # check what processes are random effects
input$random = NULL # so inner optimization won't change simulated RE
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
# Note: do.fit must be FALSE (no modeling fitting yet)

om_with_data <- update_om_fn(om, seed = 123, random = random)

assess.interval <- 2 # 
base.years      <- year_start:year_end # Burn-in period
first.year      <- head(base.years,1)
terminal.year   <- tail(base.years,1)
assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)

mods <- list() # Create a list to save MSE outputs
```


### 2. Conduct MSE

```{r, eval=FALSE}
sub.dir = "Results"
dir.create(file.path(getwd(), sub.dir), recursive = TRUE)

library(doParallel)
library(foreach)

detectCores() # check how many cores available

cluster <- makeCluster(10)
registerDoParallel(cluster)

# Define percentFXSPR for each model
percent_vec <- c(60, 70, 80, 90, 100)

# Loop over 5 models
for (model_num in seq_along(percent_vec)) {
 
  percent_fxspr <- percent_vec[model_num]
 
  cat(sprintf("Starting Model %d with percentFXSPR = %d\n", model_num, percent_fxspr))
 
  foreach(i = 1:10) %dopar% {
   
    # Wrap in tryCatch to handle errors
    tryCatch({
      library(wham)
      library(SPASAM.MSE)
     
      om_with_data <- update_om_fn(om, seed = 1010101 + i, random = random)
     
      mod <- loop_through_fn(
        om = om_with_data,
        em_info = info,
        random = random,
        NAA_re_em = NAA_re,
        em.opt = list(separate.em = FALSE, separate.em.type = 1, do.move = FALSE, est.move = FALSE),
        assess_years = assess.years,
        assess_interval = assess.interval,
        base_years = base.years,
        year.use = 30,
        hcr = list(hcr.type = 1, hcr.opts = list(use_FXSPR = TRUE, percentFXSPR = percent_fxspr)),
        seed = 1010101 + i,
        save.last.em = FALSE
      )
     
      saveRDS(mod, file.path(sub.dir, sprintf("Mod%d_%03d.RDS", model_num, i)))
     
    }, error = function(e) {
      cat(sprintf("Error in Model %d, Replicate %d: %s\n", model_num, i, e$message))
      saveRDS(NULL, file.path(sub.dir, sprintf("Mod%d_%03d_ERROR.RDS", model_num, i)))
    })
  }
}

stopCluster(cluster)
```

### 3. Collect Results
```{r, eval=TRUE}
model_nums <- 1:5
nsim <- 10 # number of simulations/seed

mods <- lapply(1:nsim, function(r) {
  
  mod_list <- lapply(model_nums, function(m) {
    file_path <- file.path("Results", sprintf("Mod%d_%03d.RDS", m, r))
    readRDS(file_path)
  })
  
  names(mod_list) <- paste0("Mod", model_nums)
  
  return(mod_list)
})
```

### 4. Plot MSE Results
```{r, eval=FALSE}
plot_mse_output(mods,
                main_dir = getwd(),
                output_dir = "Report",
                output_format = c("html"), # or html or png
                width = 10, height = 7, dpi = 300,
                col.opt = "D",
                new_model_names = c("M1","M2","M3","M4","M5"),
                base.model = "M1",
                start.years = 31,
                use.n.years.first = 5,
                use.n.years.last = 5)

```

### 5. MSE Performance Metrics
<embed src="Report/mse_report.pdf" width="110%" height="800px" type="application/pdf">